---
title: "Visualising the Heavens" 
subtitle: "An Interactive Spatio-Temporal Visualization of Asteroid Discoveries in Tucson"
author: "Data Dynamos"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
```

```{r}
#| label: load-dataset
#| message: false
```

## Overview

Creation of a interactive website for visualization of asteroids and describe individual asteroid data based on NASA's asteroid data set.

## Did You Know?

Did you know that Tucson, Arizona, is a hotspot for asteroid discoveries, with its clear skies and advanced observatories playing a key role in identifying new celestial bodies?

## Introduction

In the boundless expanse of the solar system, a hidden universe of rocky wanderers awaits for exploration. The secrets to solving the riddles surrounding the genesis and evolution of the solar system can be found within these space nomads, also known as asteroids, which are frequently written off as nothing more than stray pieces of a chaotic universe. In this project, we aim to create an interactive website to examine the asteroids from the discoveries made in Tucson Research centre.

For countless years, astrophysicists and enthusiasts have been enthralled with the fascinating universe of asteroids. These cosmic nomads captivate us with their enduring mystique; they are frequently perceived as leftovers of the birth of our solar system. This topic is brought to light by the enigmatic nature of asteroids and the mysteries they hold that have not yet been discovered. Motivated by the universe's fascination and secrets, our research team discovered a shared passion for investigating the asteroid realm. It will provide a unique opportunity to look into the unexplored regions of space and learn more about them.

## Project Overview

The website will serve as an educational and exploratory tool for both astronomy enthusiasts and the general public. It will feature interactive visualizations that allow users to explore various aspects of the asteroids discovered in Tucson, including their names, designations, discovery dates, locations, and the names of discoverers. 

## Data Utilization

The NASA data portal is the source of the asteroid data set for the project. The data set offers information about the asteroids, including their names, designation, discovery date, discovery location and the names of the individuals or institutions responsible for the discoveries. The data set retains its original presentation, without alterations to account for the loss of diacritical marks due to limitations in the ASCII character set. This version of data set includes 9999 entries and contains the information for all asteroids numbered as of 2nd February 1999 batch of Minor planet Circulars. These information were either supplied by the Minor Planet Center or taken from the Minor Planet Circular, which the organization distributed on behalf of Commission 20 of the International Astronomical Union.

We will be utilizing the following entries:

• **Catalog Number**: The unique identifier for each asteroid, allowing us to track and reference them systematically.

• **Name**: The official name of the asteroid, which can offer insights into cultural and historical contexts of the discovery.

• **Principal Provisional Designation**: The initial temporary label given to the asteroid, crucial for cross-referencing with historical records and scientific literature.

• **Discovery Date**: The date each asteroid was discovered, which will be used to create the temporal component of our visualizations.

• **Discovery Location**: The geographic location where the asteroid was first observed, a key element in mapping the discoveries.

• **Discoverer(s) Name**: Information about who discovered the asteroid, highlighting the human element in the field of astronomy.

## Goal

This project aims to create an interactive website, focusing on the visualization of data from asteroid discoveries made in Tucson, as detailed in the data set "Asteroid Names and Discovery V4.0" (DATA_SET_ID: EAR-A-5-DDR-ASTNAMES-DISCOVERY-V4.0).

Various Components of dashboard includes:

#### 1. Timeline Visualization:

An interactive timeline will showcase the dates of asteroid discoveries. Users can slide through time to see the pace of discovery and significant milestones.

#### 2. Map Integration:

Utilizing the discovery location data, we will provide a map visualization where users can zoom in on Tucson and surrounding discovery sites, providing a spatial context to the data.

#### 3. Discovery Details:

By selecting individual asteroids, users will be able to view detailed information such as the asteroid's catalog number, provisional designation, and the name(s) of the discoverer(s).

#### 4. Educational Information:

The website will also include educational resources about asteroids, their significance in our solar system, and the history of asteroid discovery in Tucson.

## Implementation Plan

#### Technical Considerations:

The website will be developed using HTML, CSS, and JavaScript. The D3.js library will be instrumental for creating the data-driven visualizations. We will ensure the site is responsive and accessible across various devices and browsers. 

#### Challenges and Limitations

A key challenge is the availability of the data set since it has been removed from NASA website. Also, the presentation of the inconsistencies in the data set, such as varying name presentations and location specifics will be challenging. We will need to decide whether to standardize these entries for consistency or to present them as they are, maintaining the data set's authenticity. Finally, the limitation of data stopping at 1999 requires a clear communication to the users that more recent discoveries will not be included.

## Plan Of Attack

-   **Week 1: (Data gathering and preparation)**

    -   Deema & Shakir - Data gathering from the NASA website for asteroids and data validation of all relevant variables like - Catalog number, name, Principal Provisional Designation, Discovery Date, Discovery Location, Discoverer(s) Name.
    -   Ansh & Rahul - Begin the initial data cleaning process.
    -   Swati - Create a shared repository or document to organize the collected data efficiently.
    -   Everyone - Document the sources and formats of the collected data sets.

-   **Week 2: (Data Exploration and Visualization)**

    -   Everyone - Analyse the cleaned data set to understand its structure.
    -   Shakir, Ansh & Deema - Begin the process of designing the initial map visualizations of asteroids' positions.
    -   Swati & Rahul - Experiment with basic visualization techniques and work on creating visual representations.

-   **Week 3: (Interactive Features Development)**

    -   Shakir, Ansh & Deema - Work on data integration for the interactive features within the dashboard visualization.
    -   Swati & Rahul - Try and implement wireframes for the interactive features.
    -   Everyone - Perform initial rounds of testing to validate the data from user perspective.
    -   Swati & Ansh - Peer review code and ensure smooth operation.
    -   Everyone - Dynamically come up with ways (ideas) to make the user experience better.

-   **Week 4: (Testing, Refinement, and Documentation)**

    -   Everyone - Address any issues or bugs identified during testing.
    -   Shakir, Swati & Ansh - Enhance the aesthetics and user experience of the map visualizations based on user feedback.
    -   Raul & Deema - Work on final testing and quality assurance to guarantee a bug-free release.
    -   Everyone - Write comprehensive guide and complete documentation for the project in GitHub.
    -   Everyone - Work on final project presentation.

## Organization

The following are the folders involved in the Project repository.

-   **'data/':** Used for storing any necessary data files for the project, such as input files.

-   **'images/':** Used for storing image files used in the project.

-   **'\_extra/':** Used to experiment on analysis which won't impact our project workflow.

-   **'\_freeze/':** This folder is used to store the generated files during the build process. These files represent the frozen state of the website at a specific point in time.

-   **'.github/':** Folder for storing github templates and workflow.

## Conclusion

This interactive website will not only fulfill the requirements for our final project but also serve as a bridge between historical astronomical data and modern-day data visualization techniques. It will contribute to the greater public understanding of Tucson's role in asteroid discovery and the importance of data visualization in narrating the story of our skies.

::: callout-note
## Note:

These are the planned approaches, and we intend to explore and solve the problem statement which we came up with. Parts of our approach might change in the final project.
:::
